Manuscript Draft

### 2019-01-04 manuscript.tex

\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Manuscript},
            pdfauthor={Sally E. Claridge, Daniel M. Charytonowicz, \& Adam A. Margolin},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Manuscript}
    \author{Sally E. Claridge\textsuperscript{1}, Daniel M.
Charytonowicz\textsuperscript{1}, \& Adam A.
Margolin\textsuperscript{1, 2}}
    \date{}

@@

We derived our clinical associations from a new project conducted by the
Variant Interpretation for Cancer Consortium (VICC;
\url{https://cancervariants.org/}), a Driver Project for the Global
Alliance for Genomics and Health (GA4GH) (Genomics \& Health, 2016). The
VICC has curated annotations of known cancer variants at varying levels
of evidence from multiple resources (A. H. Wagner et al., 2018): the
Precision Medicine Knowledgebase (L. Huang et al., 2016), MolecularMatch
(\url{https://www.molecularmatch.com/index.html}), OncoKB (Chakravarty
et al., 2017), Jackson Labs Clinical Knowledgebase (S. E. Patterson et
al., 2016), Clinical Interpretations of Variance in Cancers (CIViC)
(Griffith et al., 2017), and the Cancer Genome Interpreter (CGI)
(Tamborero et al., 2018). These harmonized variants are hosted on an
ElasticSearch (Kibana v6.0) platform called Genotype to Phenotype (G2P,
\url{https://search.cancervariants.org/\#*}) that allows users to query
and filter aggregated evidence from the various databases listed above
as well as the GA4GH beacon service
(\url{http://beacon-network.org/\#/}), which yields access to genetic
mutation data from over 200 datasets as of 2016 (Genomics \& Health,
2016). For drug screen datasets, we focus on three well-known,
large-scale drug projects: the CCLE and GDSC, which were mentioned
above, and the Center for the Science of Therapeutics at the Broad
Institute's Cancer Therapeutics Response Portal (CTRP) dataset (Basu et
al., 2013; M. G. Rees et al., 2016; B. Seashore-Ludlow et al., 2015). We
also analyzed the results from a large-scale CRISPR/Cas9 screen
conducted by the Broad Institute's Cancer Dependency Map project
(DepMap) using the Avana knockout library (Doench et al., 2016; Meyers
et al., 2017), which will address whether gene essentiality screens via
CRISPR yield results comparable to functional drug screens.

@@

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:table-g2p-nonCID}Distribution of level-A G2P drugs excluded from analyses}
\begin{tabular}{llll}
\toprule
Compound & \multicolumn{1}{c}{Compound identifier} & \multicolumn{1}{c}{Number of G2P interpretations} & \multicolumn{1}{c}{Proportion of excluded interpretations (\%)}\\
\midrule
cetuximab & CHEMBL1201577 & 74 & 18.41\\
cisplatin & CHEMBL2068237 & 3 & 0.75\\
egfr & SID160769799 & 1 & 0.25\\
ipilimumab & CHEMBL1789844 & 2 & 0.50\\
mab & CHEMBL2109423 & 16 & 3.98\\
nivolumab & CHEMBL2108738 & 4 & 1.00\\
panitumumab & CHEMBL1201827 & 75 & 18.66\\
pembrolizumab & CHEMBL3137343 & 3 & 0.75\\
pertuzumab & CHEMBL2007641 & 1 & 0.25\\
trastuzumab & CHEMBL1201585 & 9 & 2.24\\
NA & NA & 214 & 53.23\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} mAb = monoclonal antibody; NA indicates G2P interpretation does not have an associated compound}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

G2P designed a set of standards for stratifying their database's cancer
variant interpretations, ranging from preclinical data at the
low-evidence end (level D) to clinically actionable interpretations at
the high-evidence end (level A). To test only these clinically
actionable associations, we filtered the G2P dataset for only level-A
G2P interpretations, of which there were 1,296 (see Methods). All
compounds screened within the GDSC, CTRP, and CCLE datasets are
considered \enquote{small molecules} and are respectively linked to
Compound (CID) numbers used for indexing in the PubChem database
(\url{https://pubchem.ncbi.nlm.nih.gov/}). As a result, non-small
molecules, which include proteins and biologics (i.e.~monoclonal
antibodies), that had level-A G2P evidence could not be included in our
comparisons. Thus, G2P interpretations for which the compound did not
have a matching CID code were excluded from further analysis as were
interpretations that did not have an associated compound, resulting in
removal of 402 interpretations Table~\ref{tab:table-g2p-nonCID}. It
should be noted that cisplatin, a small molecule, was excluded due to
its being indexed with a ChEMBLdb identifier
(\url{https://www.ebi.ac.uk/chembl/}), which we did not manually
re-index to a CID to avoid accidental mischaracterization.

After filtering, 894 level-A G2P interpretations were carried forward in
subsequent analyses, constituting 57 unique drugs and 34 unique genes in
88 distinct level-A combinations (\emph{Figure 1A}). The number of
level-A interpretations per drug ranged from 1 to 176 (\(M =\) 15.68,
\(SD =\) 34.63), 1 to 264 (\(M =\) 26.24, \(SD =\) 61.58) per gene, and
for each level-A drug-gene combination, the range was 1 to 103 (\(M =\)
10.14, \(SD =\) 19.80). These unique drugs, genes, and combinations are
hereafter referred to as G2P drugs, G2P genes, and G2P associations. Of
the 57 unique G2P drugs, 37 of them were screened in at least one of the
drug screen datasets, 14 were screened in two of the three datasets, and
only three drugs were screened in all three datasets. The remaining 20
un-screened G2P drugs suggests an incompleteness of these drug screens,
though these projects frequently release new screening data that may
include these drugs in future releases.

Using cell-line-specific point mutation annotations from the CCLE, we
compared z-score-transformed area under the dose-response curve (AUC)
distributions between mutant and wildtype cell lines for all G2P genes
for each G2P drug tested in the CCLE, CTRP, and GDSC drug screens
(\emph{Figures 1B and 2A}). At uncorrected \(\alpha = 0.05\), Wilcoxon
rank-sum tests revealed that in CCLE, the AUC distributions of mutant
and wildtype groups for 10 out of 170 (5.88\%) tested drug-gene
combinations were significantly different, 91 out of 918 (9.91\%) were
significantly different for CTRP, and 61 out of 850 (7.18\%) for GDSC.
We observe little concordance between the number of interpretations for
each G2P association and the significance of mutation status on AUC for
said G2P association, suggesting that mutation status alone is not a
robust predictor of clinically actionable genetic targets. This led us
to examine the correlations between AUC and gene expression and AUC and
copy number, two other genomic features that are known to associate with
cancer and drug response.

@@

In the level-A G2P interpretations, there were 102 unique
drug-gene-lineage combinations (DGLs), representing seven of our curated
cancer lineages: cancer, breast cancer, leukemia, lung cancer, ovarian
cancer, skin cancer, and thyroid cancer. CCLE had data for nine DGLs,
CTRP had data for 61, and GDSC had data for 53, yielding a total of 73
unique DGLs across all three datasets.

@@

CRISPR/Cas9, CCLE mutation calls, and CCLE copy number data were all
retrieved from the Broad Institute's Cancer Dependency Map (DepMap)
Public 18Q4 release via the DepMap data portal
(\url{https://depmap.org/portal/download/}) (DepMap, 2018). Gene
expression data was retrieved from the DepMap Public 18Q3 release (Broad
DepMap, 2018) from the same data portal. For cell lines with no CCLE
annotation, the DepMap group drew raw copy number and mutation data from
whole exome sequencing data produced by the Wellcome Trust Sanger
Institute {[}Catalogue Of Somatic Mutations In Cancer (COSMIC,
\url{https://cancer.sanger.ac.uk/cell_lines}) (S. Bamford et al., 2004;
S. A. Forbes et al., 2017); European Genome-phenome Archive (Lappalainen
et al., 2015), accession number EGAD00001001039{]} and processed the
data following the CCLE pipelines to ensure consistency.

@@

Gene effect scores for 17,634 genes in 517 cell lines were inferred from
a CRISPR/Cas9 (clustered regularly interspaced short palindromic
repeats/CRISPR-associated 9) screen using the Broad Institute's Avana
knockout library (Doench et al., 2016). The Broad Institute's data
processing and screening methods are available from the figshare record
(DepMap, 2018) and the original publication of CERES, the algorithm that
computes inferred gene dependency scores (Meyers et al., 2017). The
DepMap releases new data quarterly, with the current data set being the
18Q4 release. This release has 175 more cell lines than the original
data release that was published with CERES (dataset: gene\_effect.csv,
accessed 15 November 2018).

@@

To preclude variation in genomic feature calls across the datasets, we
used identical annotations from CCLE for all cell lines screened in the
three datasets. The CCLE provided mutation annotations in 19,280 genes
across 1,596 cell lines (dataset: depmap\_18Q4\_mutation\_calls.csv,
accessed 14 November 2018). CCLE called substitutions using MuTect
(Cibulskis et al., 2013) and annotated variants using Oncotator (Ramos
et al., 2015) and indels using Indelocator
(\url{https://software.broadinstitute.org/cancer/cga/indelocator}). We
filtered mutation calls for point mutations, defined as
single-nucleotide insertions, deletions, and substitutions, regardless
of result, i.e.~frameshift or missense mutation. For this analysis, cell
lines harboring non-silent point mutations were considered
\enquote{mutant} for the gene in question. All other mutations and genes
without annotation in the Mutation Annotation Format (MAF) file were
considered \enquote{wildtype.} \enquote{Mutant} and \enquote{wildtype}
calls were binarily encoded per-gene/per-cell line, regardless of
quantity of harbored mutations per gene in a given cell line.

@@

The CCLE generated genomic copy number (CN) data using the Affymetrix
Genome-Wide Human SNP Array 6.0 and GenePattern pipeline (Network, 2008)
and normalized segmented CN log2-ratios for 23,299 genes across 1,098
cell lines using circular binary segmentation (Olshen, Venkatraman,
Lucito, \& Wigler, 2004) (dataset: public\_18Q4\_gene\_cn.csv, accessed
15 November 2018).

@@

For \emph{Figure 1}, all data were imported and manipulated using pandas
(v0.23.0) (McKinney, 2011) and plotted using seaborn (v0.90) (Waskom,
2012), which sat on top of matplotlib (v2.2.2) (Hunter, 2007) and was
executed on Python (v2.7.15) (Python Software Foundation,
\url{https://www.python.org/}) in an IPython (v5.5.0) (Pérez \& Granger,
2007) kernel within a localhost Jupyter (v5.2.2) notebook (Kluyver et
al., 2016). All other analyses were conducted in R (v3.5.0) (R
Foundation for Statistical Computing, \url{https://www.R-project.org/})
in an RStudio environment (v1.1.447) (RStudio, 2012). All other plots
were drawn using the ggplot2 (v3.0.0) R package (Wickham, 2016).

@@

Wilcoxon tests were computed using the ggpubr (v.) R package
(Kassambara, 2018).

@@

\hypertarget{ref-bairoch2018}{}
Bairoch, A. (2018). The Cellosaurus, a Cell-Line Knowledge Resource.
\emph{Journal of Biomolecular Techniques : JBT}, \emph{29}(2), 25--38.
doi:\href{https://doi.org/10.7171/jbt.18-2902-002}{10.7171/jbt.18-2902-002}

@@

\hypertarget{ref-basu2013}{}
Basu, A., Bodycombe, N. E., Cheah, J. H., Price, E. V., Liu, K.,
Schaefer, G. I., \ldots{} Schreiber, S. L. (2013). An interactive
resource to identify cancer genetic and lineage dependencies targeted by
small molecules. \emph{Cell, Cell}, \emph{154, 154}(5, 5), 1151,
1151--1161.
doi:\href{https://doi.org/10.1016/j.cell.2013.08.003,\%2010.1016/j.cell.2013.08.003}{10.1016/j.cell.2013.08.003, 10.1016/j.cell.2013.08.003}

@@

@@

\hypertarget{ref-broaddepmap2018a}{}
DepMap, B. (2018, January 11). DepMap Achilles 18Q4 public. Figshare.
Retrieved from
\url{https://figshare.com/articles/DepMap_Achilles_18Q4_public/7270880}

@@

\hypertarget{ref-domcke2013}{}
Domcke, S., Sinha, R., Levine, D. A., Sander, C., \& Schultz, N. (2013).
Evaluating cell lines as tumour models by comparison of genomic
profiles. \emph{Nature Communications}, \emph{4}.
doi:\href{https://doi.org/10.1038/ncomms3126}{10.1038/ncomms3126}

\hypertarget{ref-forbes2017}{}
Forbes, S. A., Beare, D., Boutselakis, H., Bamford, S., Bindal, N.,
Tate, J., \ldots{} Campbell, P. J. (2017). COSMIC: Somatic cancer
genetics at high-resolution. \emph{Nucleic Acids Research},
\emph{45}(D1), D777--D783.
doi:\href{https://doi.org/10.1093/nar/gkw1121}{10.1093/nar/gkw1121}

@@

\hypertarget{ref-theglobalallianceforgenomicsandhealth2016}{}
Genomics, T. G. A. for, \& Health. (2016). A federated ecosystem for
sharing genomic, clinical data. \emph{Science}, \emph{352}(6291),
1278--1280.
doi:\href{https://doi.org/10.1126/science.aaf6162}{10.1126/science.aaf6162}

@@

\hypertarget{ref-hatzis2014}{}
Hatzis, C., Bedard, P. L., Juul Birkbak, N., Beck, A. H., Aerts, H. J.
W. L., Stern, D. F., \ldots{} Haibe-Kains, B. (2014). Enhancing
Reproducibility in Cancer Drug Screening: How Do We Move Forward?
\emph{Cancer Research}, \emph{74}(15), 4016--4023.
doi:\href{https://doi.org/10.1158/0008-5472.CAN-14-0725}{10.1158/0008-5472.CAN-14-0725}

@@

\hypertarget{ref-heiser2009}{}
Heiser, L. M., Wang, N. J., Talcott, C. L., Laderoute, K. R., Knapp, M.,
Guan, Y., \ldots{} Spellman, P. T. (2009). Integrated analysis of breast
cancer cell lines reveals unique signaling pathways. \emph{Genome
Biology}, \emph{10}(3), R31.
doi:\href{https://doi.org/10.1186/gb-2009-10-3-r31}{10.1186/gb-2009-10-3-r31}

\hypertarget{ref-holbeck2010}{}
Holbeck, S. L., Collins, J. M., \& Doroshow, J. H. (2010). Analysis of
FDA-Approved Anti-Cancer Agents in the NCI60 Panel of Human Tumor Cell
Lines. \emph{Molecular Cancer Therapeutics}, \emph{9}(5), 1451--1460.
doi:\href{https://doi.org/10.1158/1535-7163.MCT-10-0106}{10.1158/1535-7163.MCT-10-0106}

\hypertarget{ref-huang2016}{}
Huang, L., Fernandes, H., Zia, H., Tavassoli, P., Rennert, H., Pisapia,
D., \ldots{} Elemento, O. (2016). The cancer precision medicine
knowledge base for structured clinical-grade mutations and
interpretations. \emph{Journal of the American Medical Informatics
Association : JAMIA}, \emph{24}(3), 513--519.
doi:\href{https://doi.org/10.1093/jamia/ocw148}{10.1093/jamia/ocw148}

@@

\hypertarget{ref-thecancergenomeatlasresearchnetwork2008}{}
Network, T. C. G. A. R. (2008). Comprehensive genomic characterization
defines human glioblastoma genes and core pathways, \emph{455},
1061--1068.
doi:\href{https://doi.org/10.1038/nature07385}{10.1038/nature07385}

@@

>>>>

\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Manuscript Draft},
            pdfauthor={Sally E. Claridge, Daniel M. Charytonowicz, \& Adam A. Margolin},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Manuscript Draft}
    \author{Sally E. Claridge\textsuperscript{1}, Daniel M.
Charytonowicz\textsuperscript{1}, \& Adam A.
Margolin\textsuperscript{1, 2}}
    \date{}

@@

We derived our clinical associations from a new project conducted by the
Variant Interpretation for Cancer Consortium (VICC;
\url{https://cancervariants.org/}), a Driver Project for the Global
Alliance for Genomics and Health (GA4GH) (The Global Alliance for
Genomics and Health, 2016). The VICC has curated annotations of known
cancer variants at varying levels of evidence from multiple resources
(A. H. Wagner et al., 2018): the Precision Medicine Knowledgebase (L.
Huang et al., 2016), MolecularMatch
(\url{https://www.molecularmatch.com/index.html}), OncoKB (Chakravarty
et al., 2017), Jackson Labs Clinical Knowledgebase (S. E. Patterson et
al., 2016), Clinical Interpretations of Variance in Cancers (CIViC)
(Griffith et al., 2017), and the Cancer Genome Interpreter (CGI)
(Tamborero et al., 2018). These harmonized variants are hosted on an
ElasticSearch (Kibana v6.0) platform called Genotype to Phenotype (G2P,
\url{https://search.cancervariants.org/\#*}) that allows users to query
and filter aggregated evidence from the various databases listed above
as well as the GA4GH beacon service
(\url{http://beacon-network.org/\#/}), which yields access to genetic
mutation data from over 200 datasets as of 2016 (The Global Alliance for
Genomics and Health, 2016). For drug screen datasets, we focus on three
well-known, large-scale drug projects: the CCLE and GDSC, which were
mentioned above, and the Center for the Science of Therapeutics at the
Broad Institute's Cancer Therapeutics Response Portal (CTRP) dataset
(Basu et al., 2013; M. G. Rees et al., 2016; B. Seashore-Ludlow et al.,
2015). We also analyzed the results from a large-scale CRISPR/Cas9
screen conducted by the Broad Institute's Cancer Dependency Map project
(DepMap) using the Avana knockout library (Doench et al., 2016; Meyers
et al., 2017), which will address whether gene essentiality screens via
CRISPR yield results comparable to functional drug screens.

@@

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:table-g2p-nonCID}Distribution of level-A G2P drugs excluded from analyses}
\begin{tabular}{llll}
\toprule
Compound & \multicolumn{1}{c}{Compound identifier} & \multicolumn{1}{c}{Number of G2P interpretations} & \multicolumn{1}{c}{Proportion of excluded interpretations (\%)}\\
\midrule
cetuximab & CHEMBL1201577 & 74 & 18.41\\
cisplatin & CHEMBL2068237 & 3 & 0.75\\
EGFR & SID160769799 & 1 & 0.25\\
ipilimumab & CHEMBL1789844 & 2 & 0.50\\
mAb & CHEMBL2109423 & 16 & 3.98\\
nivolumab & CHEMBL2108738 & 4 & 1.00\\
panitumumab & CHEMBL1201827 & 75 & 18.66\\
pembrolizumab & CHEMBL3137343 & 3 & 0.75\\
pertuzumab & CHEMBL2007641 & 1 & 0.25\\
trastuzumab & CHEMBL1201585 & 9 & 2.24\\
NA & NA & 214 & 53.23\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} mAb = monoclonal antibody; NA indicates G2P interpretation does not have an associated compound}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

G2P designed a set of standards for stratifying their database's cancer
variant interpretations, ranging from preclinical data at the
low-evidence end (level D) to clinically actionable interpretations at
the high-evidence end (level A). To test only these clinically
actionable associations, we filtered the G2P dataset for only level-A
G2P interpretations, of which there were 1,296 (see Methods). All
compounds screened within the GDSC, CTRP, and CCLE datasets are
considered \enquote{small molecules} and are respectively linked to
Compound (CID) numbers used for indexing in the PubChem database
(\url{https://pubchem.ncbi.nlm.nih.gov/}). As a result, non-small
molecules, which include proteins and biologics (i.e.~monoclonal
antibodies), that had level-A G2P evidence could not be included in our
comparisons. Thus, G2P interpretations for which the compound did not
have a matching CID code were excluded from further analysis as were
interpretations that did not have an associated compound, resulting in
removal of 402 interpretations (Table~\ref{tab:table-g2p-nonCID}). It
should be noted that cisplatin, a small molecule, was excluded due to
its being indexed with a ChEMBLdb identifier
(\url{https://www.ebi.ac.uk/chembl/}), which we did not manually
re-index to a CID to avoid accidental mischaracterization.

After filtering, 894 level-A G2P interpretations were carried forward in
subsequent analyses, constituting 57 unique drugs and 34 unique genes in
88 distinct level-A combinations (\emph{Figure 1A}). The number of
level-A interpretations per drug ranged from 1 to 176 (\(M =\) 15.68,
\(SD =\) 34.63), 1 to 264 (\(M =\) 26.24, \(SD =\) 61.58) per gene, and
for each level-A drug-gene combination, the range was 1 to 103 (\(M =\)
10.14, \(SD =\) 19.80). These unique drugs, genes, and combinations are
hereafter referred to as G2P drugs, G2P genes, and G2P associations.
Filtering our screening datasets by these clinical criteria yields
panels of potentially clinically relevant omic features. Of the 57
unique G2P drugs, 37 of them were screened in at least one of the drug
screen datasets, 17 drugs were screened in at least two of the three
datasets, and only 3 were screened in all three datasets. The remaining
20 un-screened G2P drugs suggests an incompleteness of these drug
screens, though these projects frequently release new screening data
that may include these drugs in future releases.

Using cell-line-specific point mutation annotations from the CCLE, we
compared z-score-transformed area under the dose-response curve (AUC)
distributions between mutant and wildtype cell lines for all G2P genes
for each G2P drug tested in the CCLE, CTRP, and GDSC drug screens
(\emph{Figures 1B and 2A}). At uncorrected \(\alpha = 0.05\), Wilcoxon
rank-sum tests revealed that in CCLE, the AUC distributions of mutant
and wildtype groups for 18 out of 170 (10.59\%) tested drug-gene
combinations were significantly different, 135 out of 918 (14.71\%) were
significantly different for CTRP, and 89 out of 850 (10.47\%) for GDSC.
We observe little concordance between the number of interpretations for
each G2P association and the significance of mutation status on AUC for
said G2P association, suggesting that mutation status alone is not a
robust predictor of clinically actionable genetic targets. This led us
to examine the correlations between AUC and gene expression and AUC and
copy number, two other genomic features that are known to associate with
cancer and drug response.

@@

In the level-A G2P interpretations, there were 102 unique
drug-gene-lineage combinations (DGLs), representing seven of our curated
cancer lineages: cancer, lung cancer, breast cancer, leukemia, thyroid
cancer, skin cancer, and ovarian cancer. CCLE had data for 9 DGLs, CTRP
had data for 61, and GDSC had data for 53, yielding a total of 73 unique
DGLs across all three datasets. Only 7 DGLs were screened in all three
datasets,

@@

CRISPR/Cas9, CCLE mutation calls, and CCLE copy number data were all
retrieved from the Broad Institute's Cancer Dependency Map (DepMap)
Public 18Q4 release via the DepMap data portal
(\url{https://depmap.org/portal/download/}) (Broad DepMap, 2018). Gene
expression data was retrieved from the DepMap Public 18Q3 release (Broad
DepMap, 2018) from the same data portal. For cell lines with no CCLE
annotation, the DepMap group drew raw copy number and mutation data from
whole exome sequencing data produced by the Wellcome Trust Sanger
Institute {[}Catalogue Of Somatic Mutations In Cancer (COSMIC,
\url{https://cancer.sanger.ac.uk/cell_lines}) (S. Bamford et al., 2004;
S. A. Forbes et al., 2017); European Genome-phenome Archive (Lappalainen
et al., 2015), accession number EGAD00001001039{]} and processed the
data following the CCLE pipelines to ensure consistency.

@@

Gene effect scores for 17,634 genes in 517 cell lines were inferred from
a CRISPR/Cas9 (clustered regularly interspaced short palindromic
repeats/CRISPR-associated 9) screen using the Broad Institute's Avana
knockout library (Doench et al., 2016). The Broad Institute's data
processing and screening methods are available from the figshare record
(Broad DepMap, 2018) and the original publication of CERES, the
algorithm that computes inferred gene dependency scores (Meyers et al.,
2017). The DepMap releases new data quarterly, with the current data set
being the 18Q4 release. This release has 175 more cell lines than the
original data release that was published with CERES (dataset:
gene\_effect.csv, accessed 15 November 2018).

@@

To preclude variation in genomic feature calls across the datasets, we
used identical annotations from CCLE for all cell lines screened in the
three datasets. The CCLE provided mutation annotations in 19,280 genes
across 1,596 cell lines (dataset: depmap\_18Q4\_mutation\_calls.csv,
accessed 14 November 2018). CCLE called substitutions using MuTect
(Cibulskis et al., 2013) and annotated variants using Oncotator (Ramos
et al., 2015) and indels using Indelocator
(\url{https://software.broadinstitute.org/cancer/cga/indelocator}). We
filtered mutation calls for point mutations, defined as
single-nucleotide insertions, deletions, and substitutions, regardless
of result, i.e.~frameshift, missense, or nonsense mutation. For this
analysis, cell lines harboring non-silent point mutations were
considered \enquote{mutant} for the gene in question. All other
mutations and genes without annotation in the Mutation Annotation Format
(MAF) file were considered \enquote{wildtype.} \enquote{Mutant} and
\enquote{wildtype} calls were binarily encoded per-gene/per-cell line,
regardless of quantity of harbored mutations per gene in a given cell
line.

@@

The CCLE generated genomic copy number (CN) data using the Affymetrix
Genome-Wide Human SNP Array 6.0 and GenePattern pipeline (The Cancer
Genome Atlas Research Network, 2008) and normalized segmented CN
log2-ratios for 23,299 genes across 1,098 cell lines using circular
binary segmentation (Olshen, Venkatraman, Lucito, \& Wigler, 2004)
(dataset: public\_18Q4\_gene\_cn.csv, accessed 15 November 2018).

@@

For \emph{Figure 1}, all data were imported and manipulated using pandas
(v0.23.0) (McKinney, 2011) and plotted using seaborn (v0.90) (Waskom,
2012), which sat on top of matplotlib (v2.2.2) (Hunter, 2007) and was
executed on Python (v2.7.15) (Python Software Foundation,
\url{https://www.python.org/}) in an IPython (v5.5.0) (Pérez \& Granger,
2007) kernel within a localhost Jupyter (v5.2.2) notebook (Kluyver et
al., 2016). All other analyses were conducted in R (v3.5.0) (R
Development Core Team, 2010, \url{https://www.R-project.org/}) in an
RStudio environment (v1.1.447) (RStudio, 2012). All other plots were
drawn using the ggplot2 (v3.0.0) R package (Wickham, 2016).

@@

Wilcoxon tests were computed using the ggpubr (v0.1.7.999) R package
(Kassambara, 2018).

@@

\hypertarget{ref-bairoch2018}{}
Bairoch, A. (2018). The Cellosaurus, a Cell-Line Knowledge Resource.
\emph{J Biomol Tech}, \emph{29}(2), 25--38.
doi:\href{https://doi.org/10.7171/jbt.18-2902-002}{10.7171/jbt.18-2902-002}

@@

\hypertarget{ref-basu2013}{}
Basu, A., Bodycombe, N. E., Cheah, J. H., Price, E. V., Liu, K.,
Schaefer, G. I., \ldots{} Schreiber, S. L. (2013). An interactive
resource to identify cancer genetic and lineage dependencies targeted by
small molecules. \emph{Cell}, \emph{154, 154}(5, 5), 1151, 1151--1161.
doi:\href{https://doi.org/10.1016/j.cell.2013.08.003,\%2010.1016/j.cell.2013.08.003}{10.1016/j.cell.2013.08.003, 10.1016/j.cell.2013.08.003}

@@

\hypertarget{ref-broaddepmap2018a}{}
Broad DepMap. (2018, January 11). DepMap Achilles 18Q4 public. Figshare.
Retrieved from
\url{https://figshare.com/articles/DepMap_Achilles_18Q4_public/7270880}

@@

@@

\hypertarget{ref-domcke2013}{}
Domcke, S., Sinha, R., Levine, D. A., Sander, C., \& Schultz, N. (2013).
Evaluating cell lines as tumour models by comparison of genomic
profiles. \emph{Nat Commun}, \emph{4}.
doi:\href{https://doi.org/10.1038/ncomms3126}{10.1038/ncomms3126}

\hypertarget{ref-forbes2017}{}
Forbes, S. A., Beare, D., Boutselakis, H., Bamford, S., Bindal, N.,
Tate, J., \ldots{} Campbell, P. J. (2017). COSMIC: Somatic cancer
genetics at high-resolution. \emph{Nucleic Acids Res}, \emph{45}(D1),
D777--D783.
doi:\href{https://doi.org/10.1093/nar/gkw1121}{10.1093/nar/gkw1121}

@@

@@

\hypertarget{ref-hatzis2014}{}
Hatzis, C., Bedard, P. L., Juul Birkbak, N., Beck, A. H., Aerts, H. J.
W. L., Stern, D. F., \ldots{} Haibe-Kains, B. (2014). Enhancing
Reproducibility in Cancer Drug Screening: How Do We Move Forward?
\emph{Cancer Res}, \emph{74}(15), 4016--4023.
doi:\href{https://doi.org/10.1158/0008-5472.CAN-14-0725}{10.1158/0008-5472.CAN-14-0725}

@@

\hypertarget{ref-heiser2009}{}
Heiser, L. M., Wang, N. J., Talcott, C. L., Laderoute, K. R., Knapp, M.,
Guan, Y., \ldots{} Spellman, P. T. (2009). Integrated analysis of breast
cancer cell lines reveals unique signaling pathways. \emph{Genome Biol},
\emph{10}(3), R31.
doi:\href{https://doi.org/10.1186/gb-2009-10-3-r31}{10.1186/gb-2009-10-3-r31}

\hypertarget{ref-holbeck2010}{}
Holbeck, S. L., Collins, J. M., \& Doroshow, J. H. (2010). Analysis of
FDA-Approved Anti-Cancer Agents in the NCI60 Panel of Human Tumor Cell
Lines. \emph{Mol Cancer Ther}, \emph{9}(5), 1451--1460.
doi:\href{https://doi.org/10.1158/1535-7163.MCT-10-0106}{10.1158/1535-7163.MCT-10-0106}

\hypertarget{ref-huang2016}{}
Huang, L., Fernandes, H., Zia, H., Tavassoli, P., Rennert, H., Pisapia,
D., \ldots{} Elemento, O. (2016). The cancer precision medicine
knowledge base for structured clinical-grade mutations and
interpretations. \emph{J Am Med Inform Assoc}, \emph{24}(3), 513--519.
doi:\href{https://doi.org/10.1093/jamia/ocw148}{10.1093/jamia/ocw148}

@@

@@

\hypertarget{ref-rdevelopmentcoreteam2010}{}
R Development Core Team. (2010). \emph{R: A Language and Environment for
Statistical Computing}. Vienna, Austria: R Foundation for Statistical
Computing. Retrieved from \url{https://www.R-project.org/}
